\bf Visual Inference for a Social Network Model

### 2019-01-28 ASApaper/paper-child.Rnw
starting to make my way through.......

\section{Background} \label{sec:bg}
% \st{
% \begin{enumerate}
% \item significance testing: what is it and why is it important
% \item goodness-of-fit testing: what is it and why is it important
% \item difficulty of these two for sna, why do vis inf
% \end{enumerate}
% }
When choosing and fitting statistical models, there are often three key evaluations of interest: significance testing of model parameters, model goodness-of-fit tests, and determining the power of those tests. With significance testing, the data are typically assumed to come from a simple model under the null hypothesis, and additional parameters are tested to determine whether they significantly contribute to explaining the variability in the data. In the second, the model of interest is examined to determine how well it fits the data. In the final, the ability of the hypothesis test in question to detect the difference between the null and alternative hypothesis is determined.  All three of these aspects of statistical modeling increase greatly in difficulty as the complexity of the model of interest increases. 
%The more complicated the model, the harder it is to determine fit or to include or exclude a parameter.

Some particularly complicated families of models are those designed to model network change. A \textit{network} is any set of things, such as people, computers, or neurons, that are connected in some way, through social relations, internet connection, or electrical impulses in the brain. We refer to the ``things" in the network as \textit{nodes}, or \textit{actors} in a social network, and the connections as \textit{edges}, or \textit{ties} in a social network. 
Dependencies inherent to the data make network objects particularly difficult to model. %Even more challenging is the situation 
The difficulty of modeling increases when we examine more than a single instance of a network and consider the dynamics of network change between observed instances. This type of modelling for dynamic networks is often performed on social network data, such as friendship networks among students or the spread of HIV in drug users sharing needles. %sexually active communities.% of men sleeping with other men. X ST a bit less jarring hopefully. X 
%These 
It is known that most network models lack the asymptotics required to perform many well-known goodness-of-fit tests, and the maximum likelihood estimation of parameters is so difficult that it can make significance testing difficult as well, see for example \citet{holland81}.
%\hh{is the previous sentence a quote? I'm not sure that I follow.} %\st{it's a paraphrase... shall I make a quote? Also I definitely botched a word. Hopefully makes more sense now? }

In order to circumvent some of these difficulties, we propose new methods for significance testing of parameters, goodness-of-fit testing, and power calculations %of these 
for a set of social network models: stochastic actor-oriented models for dynamic network data, first introduced by \citet{saompaper}. Specifically, we are using \textit{visual inference} in place of traditional statistical methods for social network models, such as Wald or $t$-tests for significance of parameters and in- and outdegree distribution metrics for determining goodness-of-fit. Visual inference, introduced by \citet{Bujaetal}, will allow us to look at the \textit{entire} dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network such as outdegree distribution, or a $p$-value for a parameter in the model. By using visual inference to supplement traditional statistical tests, we gain a new understanding of parameters in SAOMs and fit of SAOMs to data.%In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.

>>>>

\section{Background} \label{sec:bg}
% \st{
% \begin{enumerate}
% \item significance testing: what is it and why is it important
% \item goodness-of-fit testing: what is it and why is it important
% \item difficulty of these two for sna, why do vis inf
% \end{enumerate}
% }
When selecting and fitting statistical models, there are typiccally three key assessments of interest: significance tests of parameters, goodness-of-fit tests, and power calculations of the tests. With significance testing, the null hypothesis assumes that the data come from a simple model nested within the model of interest. We conduct significance tests of one or more additional parameters to determine if they significantly explain the variability in the data. For goodness-of-fit tests, we examine one or more models of interest to assess how well they fit the data. Finally, power tests quantify the ability of the hypothesis test to detect the difference between the null and alternative hypothesis.  All three of these aspects of statistical modeling are vital for inference. 
%The more complicated the model, the harder it is to determine fit or to include or exclude a parameter.

The more complex the model, however, the harder it is to assess significance, power, and goodness-of-fit. One particularly complicated family of models are those designed to model network change. A \textit{network} is any set of things, such as people, or computers, that are connected in some way, through social relations or the Internet. In a social network, the people are the \textit{nodes}, or \textit{actors}, and the relationships between the people are \textit{edges}, or \textit{ties}. Models for networks are particularly complex, as 
dependencies inherent to network data make them  difficult to model. %Even more challenging is the situation 
And this difficulty only increases when studying \textit{dynamic networks}, the same set of nodes and their changing relationships observed at many points in time. Trying to determine the mechanism underlying change in a network over time only makes the model more complex. Yet, researchers are often interested in modeling dynamic social networks, such as friendship networks among students or collaboration networks between researchers. % X ST a bit less jarring hopefully. X 
Even some of the simplest network models, however, lack the asymptotics required to perform traditional goodness-of-fit tests \citep{holland81}. In additional, direct maximum likelihood estimation of model parameters is frequently impossible due to the intractability of the models \citep{hummeletal}. 
%\hh{is the previous sentence a quote? I'm not sure that I follow.} %\st{it's a paraphrase... shall I make a quote? Also I definitely botched a word. Hopefully makes more sense now? }
% stop 4pm on 1/28/19
In order to circumvent some of these difficulties, we propose new methods for significance testing of parameters, goodness-of-fit testing, and power calculations %of these 
for a set of social network models: stochastic actor-oriented models for dynamic network data, first introduced by \citet{saompaper}. Specifically, we are using \textit{visual inference} in place of traditional statistical methods for social network models, such as Wald or $t$-tests for significance of parameters and in- and outdegree distribution metrics for determining goodness-of-fit. Visual inference, introduced by \citet{Bujaetal}, will allow us to look at the \textit{entire} dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network such as outdegree distribution, or a $p$-value for a parameter in the model. By using visual inference to supplement traditional statistical tests, we gain a new understanding of parameters in SAOMs and fit of SAOMs to data.%In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.
